---
title: Calculation of carbon dioxide emissions from movement trajectories
subtitle: GEO880 - Computational Movment Analysis (UZH) / Patterns and Trends in Environmental Data (ZHAW)
author: Laura Haus and Antonia Hehli
date: 07.02.2023
output: html
fig_caption: yes
format: 
  html: 
    code-fold: true
    code-tools: true
execute: 
  output: false
  warning: false
toc: true
abstract: The transport sector is an important source for carbon dioxide emission and therefore an important factor in climate change. Using movement data from the Posmo app, emissions were calculated based on three approaches with increasing complexities based on distance, speed and slope. Thereby, we used 3 different validated data sets. There are significant differences in the calculations of CO~2~ emissions depending on the complexity of the approach whereas an increased complexity lead to decreased emissions. Our study revealed that the calculation of emissions depends on many factors which are difficult to include in a holistic analysis.  
---

*Same grade requested for both students*

```{r libraries, warning = FALSE, message = FALSE}
# Preliminaries
# Import libraries
library("dplyr")
library("ggplot2")
library("gridExtra")
library("leaflet")
library("lubridate")
library("readr")
library("raster")
library("RColorBrewer")
library("sf")
library("terra")
library("tidyterra")
```

# Introduction

The transport sector is one of the major contributors to the global carbon dioxide (CO~2~) emissions (Solaymani, 2019) with road transportation responsible for around 20 percent of emissions. Almost half of these emissions are caused by passenger road vehicles (Suarez et al., 2022). Due to the strong negative impact of CO~2~ emissions on climate change and the challenges arising, they are receiving gradually more concern (Dong et al., 2022). Driving behavior like acceleration as well as terrain characteristics such as a slope have valid effects on CO~2~ emissions (Dong et al., 2022; Suarez et al., 2022; Xu et al., 2022). However, not only road traffic is accountable for the emissions in the transport sector. Trains, trams, and – when also taking production emissions into account – even bicycles cause CO~2~ emissions (BAFU, 2023; Mobitool, 2020). 
Using validated movement data collected by the Posmo app from 3 students we assess the CO~2~ emissions caused by different transportation modes. The objective of this report is to analyse how much the complexity of a calculation approach changes the results for CO~2~ emissions. Furthermore, we compare the results of our calculation with the emission values provided by Posmo and assess the differences for each transportation mode.

# Data and Methods

## Data Set

```{r define function loadPosmoData then load data and filter to only Switzerland}
# define a function to load the posmo data
loadPosmoData <- function(filepath){
    read_delim(filepath, delim = ",") |>
    
    # remove the place name
    subset(select = -c(place_name)) |>
    
    # remove NA values for mode of transport, we need a specified mode of transport
    filter(!is.na(transport_mode))
}

# load all validated Posmo track files
posmo1 <- loadPosmoData("data/posmo_tracks/posmo_1.csv") # our own data
posmo2 <- loadPosmoData("data/posmo_tracks/posmo_2.csv") # our own data
posmo3 <- loadPosmoData("data/posmo_tracks/posmo_3.csv") # Posmo data tracking pool

posmo <- full_join(posmo1, posmo2)
posmo <- full_join(posmo, posmo3)

# remove rows with NA in the coordinates
posmo <- posmo[!is.na(posmo$lon_x), ]

# convert to sf object and transform to LV95
posmo <- st_as_sf(posmo, coords = c("lon_x", "lat_y"), crs = 4326) |>
  st_transform(2056)

# extract coordinates
posmo_coord <- st_coordinates(posmo)

# create one dataframe
posmo <- cbind(posmo, posmo_coord)

# read Switzerland and the cantons
CH_boundaries <- st_read("data/swissboundaries3D/CH_boundaries.gpkg", crs = 2056)
CH_cantons <- st_read("data/swissboundaries3D/CH_cantons.gpkg", crs = 2056)

# filter points that are within Switzerland
posmo <- st_filter(posmo, CH_boundaries)
```

For the analysis of the emissions, we included Posmo data from a data pool provided in the course Computational Movement Analysis. We only considered the data sets where a validation of the means of transport determined by Posmo was conducted as a precise classification of the mode is crucial for our analysis. Therefore, only 3 data sets were included (see Figure 1). Our final data set included 111’253 datapoints with the following attributes:

-	user_id
-	datetime
-	weekday
-	transport_mode
-	lon_x
-	lat_y

The three data sets were recorded over different time periods. Data set 1 ranges from April 1^st^ to May 11^th^ 2023, data set 2 from April 6^th^ to June 11^th^ and data set 3 from April 11^th^ to June 16^th^ 2023. The temporal resolution of our tracking data mostly varies between 4 and 16 seconds.

Finally, we included the emission values provided by Posmo in our analysis. 

In order to calculate our own emissions, we have compiled emission values for various means of transport for Switzerland and Germany.

```{r plot all Posmo tracks in Switzerland, output = TRUE}
#| label: fig-firstFigure
#| fig-cap: All posmo tracks from the three data sets colored by mode of transport.
# check if it worked
ggplot() +
  geom_sf(data = CH_cantons, alpha = 0) +
  geom_sf(data = posmo, aes(color = transport_mode), size = 0.8) +
  coord_sf(datum = st_crs(2056)) +
  labs(x = "E",
       y = "N",
       title = "Posmo tracks in Switzerland") +
  guides(color = guide_legend(title = "mode of transport", )) +
  theme_minimal()
```

## Pre-Processing

```{r filter transportation modes}
# get all the transportation modes
unique(posmo$transport_mode)

# save transportation modes we need in a vector
transport_modes = c("Car", "Bus", "Train", "Tram", "Bike", "E_Bike")

# filter to only the transportation modes we need
posmo <- posmo |>
  filter(transport_mode %in% transport_modes)
```

Pre-processing was necessary for further analysis. We decided to reduce our assessment area to Switzerland. First, our conducted emission values are mostly based on Swiss and German data and mode of transports in other countries may have different emissions. Second, we conducted a slope analysis which would have been computationally too expensive to calculate for multiple countries. 

We focused on transportation modes which generate CO~2~ emissions. We do not include airplanes in our considerations as we are only going to focus on trajectories in Switzerland. Finally, we included car, bus, train, tram as well as e-bikes and bikes as they produce CO~2~ emissions when taking the production process into account.

Additionally, we removed all static points as we are interested in the emissions of moving vehicles. We thereby considered a threshold of 215m. The average velocity of a bicycle is 12.8 km/h (BFS & ARE, 2023). With an assumed average tracking cycle of 10 seconds this would correspond with a distance of around 215 meters per minute. 

```{r define functions for moving window and calculate static points, output = TRUE}
# functions to determine moving window
movingWindowLag <- function(E, N, n){
  sqrt((lag(E, n) - E)^2 + (lag(N, n) - N)^2)
}

movingWindowLead <- function(E, N, n){
  sqrt((E - lead(E, n))^2 + (N - lead(N, n))^2)
}

# calculate moving window
posmo <- posmo |>
  mutate(
    minus3 = movingWindowLag(X, Y, 3),
    minus2 = movingWindowLag(X, Y, 2),
    minus1 = movingWindowLag(X, Y, 1),
    plus1 = movingWindowLead(X, Y, 1),
    plus2 = movingWindowLead(X, Y, 2),
    plus3 = movingWindowLead(X, Y, 3)
  )

# calculate the mean step length
posmo <- posmo |>
  rowwise() |>
  mutate(
    stepMean = mean(c(minus3, minus2, minus1, plus1, plus2, plus3))
  ) |>
  ungroup()

# define a column 'static'
posmo <- posmo |>
    ungroup() |>
    mutate(static = stepMean < 215)


# plot whether trajectories are static or not

#| label: fig-secondFigure
#| fig-cap: Static and non-static parts of the Posmo trajectories.

posmo |>
  ggplot(aes(X,Y)) +
  geom_point(aes(colour = static)) +
  coord_fixed()+
  labs(x = "E",
       y = "N",
       title = "Static and non-static points in the trajectories",
       color = "Static")+
  theme_minimal()

# remove static points
posmo <- posmo |>
    filter(!static)
```

We filtered velocities exceeding 200 km/h, which is the highest velocities InterCity trains in Switzerland travel (Knupfer, 2014) and set the minimum speed to 10 as this corresponds with the average velocity of bicycles. Furthermore, we excluded all slopes which are steeper than 25%, which corresponds to 14.4° being the steepest mountain pass slope in Switzerland (Wikipedia, 2022).

```{r read DHM convert Posmo data to LV03 and extract height information of DHM}
# read the raster data using the terra package
dhm25 <- terra::rast("data/DHM25/dhm25_grid_raster.asc")

# set the raster crs to LV03
crs(dhm25) <- terra::crs("+init=EPSG:21781")

# convert Posmo data to LV03 to match the CRS of the DHM25
posmo_lv03 <- posmo |>
  st_transform(crs = 21781) |>
  subset(select = -c(X, Y))

# extract the height information for each Posmo location
height_dhm <- extract(dhm25, posmo_lv03)

# join the height back to the posmo data
posmo$height <- height_dhm$dhm25_grid_raster
```

# Methods

To calculate the emissions we first had to determine values for the CO~2~ emissions for the considered transportation means. As they strongly vary depending on the sources we included multiple values for some of the modes of transport. In order to determine the influence of the complexity of an approach we also considered other factors such as the effect of velocity or slope on the CO~2~ emissions. Thus, we pursued three different approaches: a simple one, where we multiplied the average emission value by the distance; a speed approach and a slope approach. To be able to compare the results we calculated the 7-days average due to the varying time periods of the data sets. 

To assess the influence of increasing complexity for the emission values we focused our analysis solely on car emissions. 

For the speed approach we used information about the change in car emission for velocities between 100 and 130 km/h by the Umweltbundesamt (n.d.). To estimate emissions at lower velocities we used a non-linear least squares regression. We assumed that the reduction in CO~2~ emissions decreases at lower velocities and therefore set a threshold at 75 km/h at which CO~2~ emissions are constant (Metcalfe & Boulter, 2022).

To calculate the increase in CO~2~ emission due to inclination, we used the reference values from Xu et al. (2020). We fitted a linear model that calculated the percentage of increase in emissions for a given slope for each 10 km/h. With these linear models, we calculated the increase in emission for our Posmo data points based on the calculated slope. We only considered an increase in emissions when passing a positive slope. Potential emission reductions when driving downhill were not taken into account.

```{r create dataframe with constants for emissions}
# for cars: we have various sources
c_car <- c(130, 162, 122.3, 200) 

# for buses
c_bus <- c(42, 108, 88.7, 25.15)

# for trains
c_train <- c(93, 110.9)

# for trams
c_tram <- c(80, 37)

# for e-bikes
c_ebike <- 14

# for bikes
c_bike <- 8

# create a look-up table for the different emission constants
constants_emission <- data.frame(
  transport_mode = transport_modes,
  stringsAsFactors = FALSE
)

# include values for emissions and calculate mean, median, minimum and maximum
constants_emission$constant_gkm <- list(c_car, c_bus, c_train, c_tram, c_bike, c_ebike) 
constants_emission <- constants_emission |>
  mutate(
    avg = sapply(constant_gkm, mean),
    med = sapply(constant_gkm, median),
    min = sapply(constant_gkm, min),
    max = sapply(constant_gkm, max)
  )
```

*brauchen wir get_days function oder eigentlich nur get_weeks?? könnte man Funktion umschreiben?* Wir bräuchten nur get_weeks. Habe es versucht, hat auf die schnelle nicht geklappt. 

```{r define function to get the number of days per user and calculate it}
# define a function to get the number of weekdays recorded per user id
get_days <- function(user, posmo_data){
  # filter the user
  posmo_user <- posmo_data |>
    filter(user_id == user)
  
  # convert the datetime format to days and get the unique days
  recorded_days <- unique(as.Date(posmo_user$datetime, format = "%d"))
  
  # return the number of unique days
  return(length(recorded_days))
}

# save the unique user_ids in the data set
posmo_userids <- unique(posmo$user_id) 

# make an empty data frame with the number of days per users
posmo_days_per_user <- data.frame(user_id = posmo_userids)

# create an empty vector to append
number_of_days = c()
number_of_weeks = c()

# use the created function
for (i in posmo_userids){
  day = get_days(i, posmo)
  number_of_days <- append(number_of_days, day)
  week = day/7
  number_of_weeks <- append(number_of_weeks, week)
}

# add the number of days and weeks rounded to 3 digits to the user_id
posmo_days_per_user$days <- number_of_days
posmo_days_per_user$weeks <- round(number_of_weeks, 3)
```

```{r calculate emissions}
# calculate the distances [m], the timelag [s], the speed [m/s], and the speed [km/h]: use group_by to make sure it only calculates for one user id and not between different user ids
posmo <- posmo |>
  group_by(user_id)|>
  mutate(
    distance_m = sqrt((lead(X, 1) - X)^2 + (lead(Y, 1) - Y)^2),
    distance_km = distance_m/1000,
    timelag_s = as.integer(difftime(lead(datetime), datetime),
                           units = "secs"),
    speed_ms = (distance_m/timelag_s),
    speed_kmh = (speed_ms*3.6),
    speed_change_ms = lead(speed_ms, 1) - speed_ms,
    acceleration_ms2 = speed_change_ms/timelag_s,
    acceleration = acceleration_ms2 >=0
    ) |>
  ungroup()

# remove NA distance values
posmo <- posmo |> 
  filter(!is.na(distance_km))

# group by user_id and transport_mode and sum up the distance in km
posmo_traveldistance <- 
  posmo |> 
  group_by(user_id, transport_mode) |>
  summarise(sum_km = sum(distance_km))

# join the lookup table to the initial data frame
posmo_traveldistance <- posmo_traveldistance |> 
  left_join(constants_emission, by = "transport_mode")

# join the number of recording days to the data frame
posmo_traveldistance <- posmo_traveldistance |> 
  left_join(posmo_days_per_user, by = "user_id")

# multiply the CO2 emissions constant with the traveled distance and divide by the number of recorded days
posmo_traveldistance <- posmo_traveldistance |>
  mutate(co2_emissions_g = sum_km * avg,
         co2_emissions_g_per_d = co2_emissions_g/days,
         co2_emissions_g_per_w = co2_emissions_g/weeks)

# calculate the total emissions by user ID and transportation mode
emissions <- aggregate(co2_emissions_g_per_w ~ user_id + transport_mode, posmo_traveldistance, sum)
```

```{r remove speed outliers and save known speed and corresponding emission values}
# remove NA and infinitive values
posmo <- posmo |>
  filter(!is.na(speed_kmh)) |>
  filter(!is.infinite(speed_kmh))

# check range of speed values
range(posmo$speed_kmh)

posmo <- posmo |>
  filter(speed_kmh >= 10) |>
  filter(speed_kmh <= 200)

# save speed and their respective emission values [g/km] as variables
speed <- c(100, 110, 120, 130)
emission <- c(146, 158, 172, 190)
```

```{r interpolation of car speed values}
# fit a non-linear least squares regression
fit_emission <- nls(emission ~ a * exp(b * speed), start = list(a = 100, b = 0.02))

# predicting new estimation values
speed_new <- seq(30, 130, by = 5)
emission_pred_gkm <- predict(fit_emission, newdata = data.frame(speed = speed_new))

# save the new values in a data frame
est_emission_car_speed <- data.frame(speed_new, emission_pred_gkm)
```

```{r plot known and interpolated emission values depending on speed, output = TRUE}
#| label: fig-thirdFigure
#| fig-cap: Estimated car emission values based on speed. Values in red from Umweltbundesamt (n.d.).
# plot the predicted as well as the old values
ggplot()+
  geom_point(aes(x = speed_new, y = emission_pred_gkm), color = "blue")+
  geom_line(aes(x = speed_new, y = emission_pred_gkm), color = "blue")+
  geom_point(aes(x = speed, y = emission), color = "red")+
  geom_line(aes(x = speed, y = emission), color = "red")+
  labs(title = expression("Estimated CO" ["2"]~ "Emission Values for Cars Depending on Speed"),
       subtitle = "Estimated values are in blue, known values are in red.",
       x = "Speed [km/h]",
       y = expression("CO"["2"] ~ "Emission [g/km]"))+
  theme_minimal()
```

```{r include cars only join emission values and calculate emissions}
# filter to all cars
posmo_car <- posmo |>
  group_by(user_id) |>
  filter(transport_mode == "Car") |>
  mutate(speed_kmh_floored = floor(speed_kmh/5) * 5) |>
  ungroup()

# join the emission values for fast movement (car speed over 75km/h)
posmo_car_fast <- posmo_car |> 
  group_by(user_id)|>
  filter(speed_kmh_floored > 75) |>
  left_join(est_emission_car_speed, by = join_by(speed_kmh_floored == speed_new)) |>
  ungroup()

# set the emission value for slow movement (car speed under 75km/h)
posmo_car_slow <- posmo_car |>
  group_by(user_id)|>
  filter(speed_kmh_floored <= 75) |>
  mutate(emission_pred_gkm = 116.41017) |>
  ungroup()

# combine everything together
posmo_car <- rbind(posmo_car_fast, posmo_car_slow)

# group by user_id to sum up the distance in km
posmo_car_traveldistance <- 
  posmo_car |> 
  group_by(user_id, emission_pred_gkm) |>
  summarise(sum_km = sum(distance_km)) |>
  mutate(emissions_g = sum_km * emission_pred_gkm)

# sum up per user
posmo_car_emissions <- posmo_car_traveldistance |>
  group_by(user_id)|>
  summarise(sum_emissions_g = sum(emissions_g, na.rm = TRUE))

# join with the number of days & weeks per user
posmo_car_emissions <- posmo_car_emissions |> 
  left_join(posmo_days_per_user)

# calculate emissions per week
posmo_car_emissions <- posmo_car_emissions |>
  group_by(user_id)|>
  mutate(co2_emissions_g_per_w = sum_emissions_g/weeks)

# calculate the total emissions by user ID and transportation mode
emissions_car_speed <- aggregate(co2_emissions_g_per_w ~ user_id, posmo_car_emissions, sum)
```

```{r plot height values for all posmo data, output = TRUE}
#| label: fig-fourthFigure
#| fig-cap: Elevation above sea level for posmo trajectories.
ggplot() + 
  geom_sf(data = CH_cantons, aes(alpha = 100)) +
  geom_sf(data = posmo, aes(color = height)) +
  coord_sf(datum = st_crs(2056))+
  labs(x = "E",
       y = "N",
       title = "Elevation of Posmo Locations",
       subtitle = "Posmo data from Switzerland") +
  scale_color_continuous(name = expression("Elevation [m a.s.l.]"), type = "viridis") +
  scale_alpha_continuous(guide = "none") +
  theme_minimal()
```

```{r calculate slope in percentage and remove outliers}
# for all posmo data
posmo_car <- posmo_car |>
  group_by(user_id)|>
  mutate(
    height_diff = lead(height, 1) - height,
    horizontal_diff = sqrt((plus1**2)-(height_diff**2)),
    slope_percent = (height_diff / horizontal_diff)*100
    )|>
  ungroup()

# remove all NA values for slope
posmo_car <- posmo_car |> 
  filter(!is.na(slope_percent))

# print the range of the slope values in percentage
range(posmo_car$slope_percent)

# filter absolute slope values greater than 100%
posmo_car <- posmo_car |> 
  filter(abs(slope_percent) <= 25)

# check if it worked
range(posmo_car$slope_percent)
```

```{r load and mutate slope emissions, message = FALSE}
emissions_slope <- read_delim("data/Emission_Slope.csv", delim = ";") |>
  # calculate the absolute increase in carbon emission
  mutate(increase_gkm = Emission_slope_gkm - Emission_flat_gkm,
         # calculate the relative increase in carbon emission
         increase_perc = (increase_gkm/Emission_flat_gkm) * 100)

# get the range of speed in the emission file
range(emissions_slope$Speed_kmh)
```

```{r calculate increase in emission depending on gradient and speed}
filterSpeed <- function(dataset, min_speed, max_speed){
  dataset |> 
    filter(Speed_kmh >= min_speed & Speed_kmh <= max_speed)
}

# filter to the three different groups
emissions_slope_slow <- filterSpeed(emissions_slope, 40, 60)
emissions_slope_medium <- filterSpeed(emissions_slope, 70, 90)
emissions_slope_fast <- filterSpeed(emissions_slope, 100, 120)

slopeEmission <- function(slope_data){
  slope_data |> 
    group_by(Speed_kmh, Gradient) |>
    mutate(mean_increase_perc = mean(increase_perc))
}

emissions_slope_slow <- slopeEmission(emissions_slope_slow)
emissions_slope_medium <- slopeEmission(emissions_slope_medium)
emissions_slope_fast <- slopeEmission(emissions_slope_fast)

# combine all into one dataframe
emissions_slope <- emissions_slope_slow |>
  rbind(emissions_slope_medium, emissions_slope_fast) |>
  filter(Vehicle != "Large") |>
  subset(select = -c(Vehicle))
```

```{r extrapolate emission values depending on slope and speed}
filterSlope <- function(emission_data, speed_value){
  emission_data |>
    filter(Speed_kmh == speed_value)
}

# run the function for each speed
slope_50kmh <- filterSlope(emissions_slope, 50)
slope_60kmh <- filterSlope(emissions_slope, 60)
slope_70kmh <- filterSlope(emissions_slope, 70)
slope_80kmh <- filterSlope(emissions_slope, 80)
slope_90kmh <- filterSlope(emissions_slope, 90)
slope_100kmh <- filterSlope(emissions_slope, 100)
slope_120kmh <- filterSlope(emissions_slope, 120)

lm_slope50 <- lm(slope_50kmh$Gradient ~ slope_50kmh$mean_increase_perc)
lm_slope60 <- lm(slope_60kmh$Gradient ~ slope_60kmh$mean_increase_perc)
lm_slope70 <- lm(slope_70kmh$Gradient ~ slope_70kmh$mean_increase_perc)
lm_slope80 <- lm(slope_80kmh$Gradient ~ slope_80kmh$mean_increase_perc)
lm_slope90 <- lm(slope_90kmh$Gradient ~ slope_90kmh$mean_increase_perc)
lm_slope100 <- lm(slope_100kmh$Gradient ~ slope_100kmh$mean_increase_perc)
lm_slope120 <- lm(slope_120kmh$Gradient ~ slope_120kmh$mean_increase_perc)


# create data frame with coefficients for each speed level
extrapolation_slope_emissions <- data.frame(speed_kmh = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130,
                                                          140, 150, 190),
                                            intercept_q = c(0, 0, 0, 0, 
                                                            lm_slope50$coefficients[1], lm_slope60$coefficients[1],
                                                            lm_slope70$coefficients[1], lm_slope80$coefficients[1],
                                                            lm_slope90$coefficients[1],lm_slope100$coefficients[1],
                                                            (lm_slope100$coefficients[1] +
                                                               lm_slope120$coefficients[1])/2,
                                                            lm_slope120$coefficients[1], 
                                                            lm_slope120$coefficients[1], lm_slope120$coefficients[1],
                                                            lm_slope120$coefficients[1], lm_slope120$coefficients[1]),
                                            slope_m = c(1, 1, 1, 1, 
                                                        lm_slope50$coefficients[2], lm_slope60$coefficients[2],
                                                        lm_slope70$coefficients[2], lm_slope80$coefficients[2],
                                                        lm_slope90$coefficients[2],lm_slope100$coefficients[2],
                                                        (lm_slope100$coefficients[2] + lm_slope120$coefficients[2])/2,
                                                        lm_slope120$coefficients[2], lm_slope120$coefficients[2],
                                                        lm_slope120$coefficients[2], lm_slope120$coefficients[2],
                                                        lm_slope120$coefficients[2]))

# round speed values to the nearest 10 to join with extrapolation
posmo_car <- posmo_car |> 
  mutate(speed_kmh_round = round(speed_kmh, -1))

# join the extrapolation of the emission values with the posmo car data
posmo_car <- posmo_car |> 
  full_join(extrapolation_slope_emissions, by = c("speed_kmh_round" = "speed_kmh"))
```

```{r calculate car emission values depending on slope and speed}
# calculate the increase in emission for each data point
posmo_car <- posmo_car |>
  filter(slope_percent >=0) |>
  mutate(emission_increase_perc = slope_m * slope_percent + intercept_q) |>
  filter(emission_increase_perc >= 0)

# count the number of observations per user
observations_user <- posmo_car |> 
  count(user_id)

posmo_slope_car_increase <- posmo_car|>
  group_by(user_id) |> 
  summarise(sum_increase_perc = sum(emission_increase_perc)) |>
  ungroup()

# add the number of observations to the data frame with the slope emissions
posmo_slope_car_increase$observations <- observations_user$n

# get the mean increase for each person's car emission when considering the slope
posmo_slope_car_increase <- posmo_slope_car_increase|>
  group_by(user_id) |> 
  mutate(mean_increase_perc = sum_increase_perc/observations) |> 
  st_drop_geometry()

# filter the simple emission values to only "Car" values
emissions_car_simple <- emissions |> 
  filter(transport_mode == "Car")

# join the car emissions with the increase of emission because of slope
posmo_car_emissions_simple <- emissions_car_simple |> 
  left_join(posmo_slope_car_increase, by = "user_id")

emissions_car_slope <- posmo_car_emissions_simple |>
  mutate(co2_emissions_g_per_w_slope = co2_emissions_g_per_w * (1 + mean_increase_perc/100))
```

# Results

The analysis of the weekly emissions of each user shows that for all three users the car and the train provide the highest emissions whereas the emissions from the bike and the e-bike are almost negligible. Buses and trams also produce only small amounts of emissions. 
Considering the weekly distance user 1 covers the most kilometers mostly by car and train followed by user 2 who makes greater use of the bus and the bicycle.

```{r plot weekly emissions for simple approach, output = TRUE}
emissions_bar_simple <- ggplot(emissions, aes(x = user_id, y = co2_emissions_g_per_w, fill = transport_mode)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "User", 
       y = expression("CO"["2"]~ "Emissions [g/week]"),
       fill = "Transport mode") +
  labs(subtitle = "Car emissions calcualted based on distance")+
  ggtitle("Average Weeky Emissions per Transportation Mode & User") +
  theme_minimal()+
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 8)) +
  ylim(0, 48000)+
  theme(axis.text.x = element_blank())

#| label: fig-sixthFigure
#| fig-cap: CO~2~ emission based on distances per user normalized over 7 days for different modes of transport.
emissions_bar_simple
```

```{r distance per user and mode of transport, output = TRUE}
posmo <- posmo |>
  left_join(posmo_days_per_user, by = "user_id") |>
  mutate(
    weekly_distance = distance_km / weeks
  )

#| label: fig-seventhFigure
#| fig-cap: Overview over travelled distance per user and mode of transport.
ggplot(posmo, aes(x = user_id, y = weekly_distance, fill = transport_mode))+
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "User", 
       y = expression("Distance [km]"),
       fill = "Transport mode") +
  ggtitle("Weekly Distance Travelled per User and Mode of Transport") +
  theme_minimal()+
  theme(axis.text.x = element_blank())
```

*leider nicht sehr viel aussagend, evtl. emissionen für nur einen weg darstellen und dafür etw. genauer analysieren*

```{r plot weekly emissions for simple approach on map, output = TRUE}
emissions_plot_simple <- ggplot() + 
  geom_sf(data = CH_cantons, aes(alpha = 0)) +
  geom_sf(data = posmo_traveldistance, aes(color = co2_emissions_g_per_w))+
  labs(title = "Average Emissions per Mode of Transport and User",
     x = "E",
     y = "N")+
  scale_color_continuous(name = expression("CO"["2"]~ "Emissions [g/week]"), type = "viridis")+
  scale_alpha_continuous(guide = "none") +
  coord_sf(datum = st_crs(2056))+
  theme_minimal()+
  theme(plot.title = element_text(size = 14))


#| label: fig-eigthFigure
#| fig-cap: Spatial reference for CO~2~ emission based on distances per user normalized over 7 days for all modes of transport.
emissions_plot_simple
```

## Speed scenario

```{r, define segments and plot one}
rle_id <- function(vec) {
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times = x))
}

car_speed <- posmo_car |>
  left_join(constants_emission, by = "transport_mode") |>
  mutate(
    speed_emission = avg * emission_increase_perc,
    segment_id = rle_id(static)
  )
```

```{r}
car_speed_day <- car_speed |>
  filter(as.Date(car_speed$datetime) == "2023-04-15")

ggplot() + 
  #geom_sf(data = CH_cantons, aes(alpha = 0)) +
  geom_sf(data = filter(car_speed_day, car_speed_day$user_id == "6e119464-37a3-4d4e-9eaf-ef1b7e137b60"), aes(color = speed_emission))+
  labs(title = "Average Emissions per Mode of Transport and User",
     x = "E",
     y = "N")+
  scale_color_continuous(name = expression("CO"["2"]~ "Emissions [g/week]"), type = "viridis")+
  scale_alpha_continuous(guide = "none") +
  coord_sf(datum = st_crs(2056))+
  theme_minimal()+
  theme(plot.title = element_text(size = 14))
```

```{r}
ggplot() + 
  geom_sf(data = CH_cantons, aes(alpha = 0)) +
  geom_sf(data = car_speed, aes(color = speed_emission))+
  labs(title = "Average Emissions per Mode of Transport and User",
     x = "E",
     y = "N")+
  scale_color_continuous(name = expression("CO"["2"]~ "Emissions [g/week]"), type = "viridis")+
  scale_alpha_continuous(guide = "none") +
  coord_sf(datum = st_crs(2056))+
  theme_minimal()+
  theme(plot.title = element_text(size = 14))
```

```{r prepare data to enable comparison with Posmo emission data}
# add a column mode of transport to the calculated car emissions using speed information as well
emissions_car_speed <- emissions_car_speed |>
  mutate(transport_mode = "Car")

emissions_new <- emissions |>
  # remove the 'old' emission values for car
  filter(transport_mode != "Car") |>
  # replace these values with the new car emissions calculated based on speed
  rbind(emissions_car_speed)
```

```{r add car emissions calculated using slope to emission data and create bar chart, output = TRUE}
# add a column mode of transport to the calculated car emissions using speed information as well
emissions_car_slope <- emissions_car_slope |>
  mutate(transport_mode = "Car") |>
  subset(select = c(user_id, co2_emissions_g_per_w_slope, transport_mode)) |>
  st_drop_geometry() |>
  rename(co2_emissions_g_per_w = co2_emissions_g_per_w_slope)

emissions_new_slope <- emissions |>
  # remove the 'old' emission values for car
  filter(transport_mode != "Car") |>
  # replace these values with the new car emissions calculated based on speed
  rbind(emissions_car_slope)
```

```{r plot locations with emission increase due to slope, output = TRUE}
#| label: fig-ninethFigure
#| fig-cap: Spatial reference of increase in car CO~2~ emission due to slope. For better visibility, increase values are limited to values between 0.5 and 10%.
posmo_car |> 
  # filter increase values between 0.5 and 10%
  filter(emission_increase_perc >= 0.5 & emission_increase_perc <= 10) |>
  ggplot()+
  geom_sf(aes(color = emission_increase_perc))+
  geom_sf(data = CH_cantons, alpha = 0) +
  labs(title = "Increase in emission due to slope",
       subtitle = "Movement trajectories in Switzerland",
       x = "E",
       y = "N")+
  scale_color_continuous(name = expression("CO"["2"]~ "emission increase [%]"), type = "viridis")+
  coord_sf(datum = st_crs(2056))+
  theme_minimal()+
  theme(plot.title = element_text(size = 14))
```

When comparing the emissions calculated with the simple and the speed approach it is evident that the speed approach resulted in lower emission values for all users. The slope approach also calculated much smaller emission values than the simple approach. However, the values for the slope and the speed approach are quite similar (Figure ??).

```{r plot grouped bar chart for all car emission scenarios, output = TRUE}
emissions_car_simple$scenario <- "simple"
emissions_car_speed$scenario <- "speed"
emissions_car_slope$scenario <- "slope"

emissions_car_all_scenarios <- data.frame(user_id = c(emissions_car_simple$user_id, emissions_car_slope$user_id,
                                                      emissions_car_speed$user_id),
                                          co2_emissions_g_per_w = c(emissions_car_simple$co2_emissions_g_per_w,
                                                                    emissions_car_speed$co2_emissions_g_per_w,
                                                                    emissions_car_slope$co2_emissions_g_per_w),
                                          scenario = c(emissions_car_simple$scenario,emissions_car_speed$scenario, 
                                                       emissions_car_slope$scenario))

emissions_car_all_scenarios_bar <- ggplot(emissions_car_all_scenarios, aes(x = user_id, y = co2_emissions_g_per_w, fill = scenario)) + 
  geom_bar(stat = "identity", position = "dodge")+
  labs(x = "Users", 
       y = expression("CO"["2"]~ "Emissions [g/week]"),
       fill = "Approach",
       subtitle = "Car emissions calculated using different approaches") +
  ggtitle("Overall Car Emissions for All Approaches") +
  geom_text(aes(label = round(co2_emissions_g_per_w, 1)), position = position_dodge(width = 0.9), size = 3, vjust = -1.5) +
  ylim(0, 25000)+
  theme_minimal()+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))

#| label: fig-tenthFigure
#| fig-cap: CO~2~ emission values for all emission scenarios of car emissions per user and week.
emissions_car_all_scenarios_bar
```

```{r calculate daily emissions}
# define a function to get the daily emissions
dailyEmission <- function(posmo_data, constants){
  # make sure Date is provided as date and not as datetime
  posmo_data$day <- as.Date(posmo_data$datetime)
  daily_distance <- aggregate(distance_km ~ user_id + day + transport_mode, posmo_data, FUN = sum)
  daily_emission <- daily_distance |>
    left_join(constants, by = "transport_mode")
  daily_emission <- daily_emission |>
    mutate(
      avg_emission_kg = distance_km * avg / 1000,
      min_emission_kg = distance_km * min / 1000,
      max_emission_kg = distance_km * max / 1000,
    )
  return(daily_emission)
}

daily_emission <- dailyEmission(posmo, constants_emission)

emission_avg <- aggregate(avg_emission_kg ~ user_id + transport_mode, daily_emission, sum)
emission_min <- aggregate(min_emission_kg ~ user_id + transport_mode, daily_emission, sum)
emission_max <- aggregate(max_emission_kg ~ user_id + transport_mode, daily_emission, sum)

emission_CH_mode <- left_join(emission_avg, emission_min, by = c("user_id", "transport_mode")) |>
  left_join(emission_max, by = c("user_id", "transport_mode"))
```

```{r read in emission values from Posmo, message = FALSE}
posmo_emission_1 <- read_delim("data/Emission_Posmo_1.csv", delim = ";") |>
  mutate(
    user_id = "1af6fc8e-273d-4889-894b-5dadadd70a01",
    Date = dmy(Date)
  )

posmo_emission_2 <- read_delim("data/Emission_Posmo_2.csv", delim = ";") |>
  mutate(
    user_id = "6e119464-37a3-4d4e-9eaf-ef1b7e137b60",
    Date = dmy(Date)
  )

posmo_emission <- rbind(posmo_emission_1, posmo_emission_2)

emission_posmo_mode <- posmo_emission |>
  tidyr::gather(transport_mode, emission, -user_id, -Date, -Notes, -Sum_all, -Sum_CH, -Airplane) |>
  mutate(origin = "posmo") |>
  dplyr::select(user_id, Date, transport_mode, emission, origin) |>
  aggregate(emission ~ user_id + transport_mode, sum)

emission_cal_posmo <- left_join(emission_posmo_mode, emission_CH_mode, by = c("user_id", "transport_mode"))
```

The bars in Figure *??* show the uncertainty which arises when calculating emission values for different transportation modes. The ranges of emission values are especially high for cars as well as buses which is visible in the increased length of the boxes which indicate the range of the possible emissions. 

The calculated emissions from Posmo and our own calculated emission vary greatly. Especially the calculated car and the train emissions show huge differences. For the car emissions Posmo calculates values which are higher than our values. Even the highest emission values we found in the literature are still rather small compared to the Posmo emissions. In contrast, the train emissions calculated by Posmo are a lot smaller than our own emission values.

```{r, plot comparison between simple approach emission calculations and Posmo emissions, output = TRUE}
#| label: fig-eleventhFigure
#| fig-cap: Comparison of our own calculated emissions including a range of possible emissions and the Posmo emissions.
ggplot(data = emission_cal_posmo)+
  geom_linerange(aes(x = transport_mode, ymin = min_emission_kg, ymax = max_emission_kg, color = user_id), 
                 size = 5, alpha = 0.4, position = position_dodge(width = 0.8)) +
  geom_linerange(aes(x = transport_mode, ymin = avg_emission_kg-0.5, ymax = avg_emission_kg+0.5, color = user_id),
                 size = 5, position = position_dodge(width = 0.8)) +
  scale_color_manual(values = c("darkblue", "darkgreen"), 
                     labels = c("User 1", "User 2"),
                     guide = guide_legend(title = "Users")) +
  geom_point(aes(x = transport_mode, y = emission, color = user_id),
             size = 3, position = position_dodge(width = 0.8), shape = 4) +
  coord_flip() +
  labs(title = "Comparison of our own calculated and the Posmo emissions",
       x = "Mode of Transport",
       y = expression("CO"["2"] ~ "Emission [g/km]"))
```

```{r}
posmo_emission_week <- emission_posmo_mode |>
  filter(transport_mode == "Car") |>
  left_join(posmo_days_per_user, by = "user_id") |>
  mutate(
    posmo_emission = emission/weeks
  ) |>
  left_join(constants_emission, by = "transport_mode") |> 
  left_join(emissions_car_simple, by = c("user_id", "transport_mode")) |>
  mutate(
    simple_avg = co2_emissions_g_per_w / 1000,
    simple_max = simple_avg / avg * max,
    simple_min = simple_avg / avg * min
  ) |>
  left_join(emissions_car_slope, by = c("user_id", "transport_mode")) |>
  mutate(
    slope_emission = co2_emissions_g_per_w.y / 1000
  ) |>
  left_join(emissions_car_speed, by = c("user_id", "transport_mode")) |>
  mutate(
    speed_emission = co2_emissions_g_per_w / 1000
  ) |>
  subset(select = c("user_id", "posmo_emission", "simple_avg", "simple_max", "simple_min", "slope_emission", "speed_emission"))
```

Comparing the overall weekly emission values for the different approaches with the Posmo emissions reveals that our overall emissions are lower than the one from Posmo. Only with very extreme values for the simple approach values similar to the Posmo one can be achieved. 

```{r, plot comparison weekly emissions between all three approaches, output = TRUE}
#| label: fig-twelfthFigure
#| fig-cap: Comparison of our own calculated emissions including a range of possible emissions and the Posmo emissions.
ggplot(data = posmo_emission_week)+
  geom_linerange(aes(x = user_id, ymin = simple_min, ymax = simple_max), 
                 size = 5, alpha = 0.4, position = position_dodge(width = 0.8), color = "darkblue") +
  geom_linerange(aes(x = user_id, ymin = simple_avg-0.1, ymax = simple_avg+0.1),
                 size = 5, position = position_dodge(width = 0.8), color = "darkblue") +
  geom_point(aes(x = user_id, y = posmo_emission),
             size = 3, position = position_dodge(width = 0.8), shape = 4, color = "red") +
  geom_point(aes(x = user_id, y = speed_emission),
             size = 3, position = position_dodge(width = 0.8), shape = 4, color = "green") +
  geom_point(aes(x = user_id, y = slope_emission),
             size = 3, position = position_dodge(width = 0.8), shape = 4, color = "blue") +
  coord_flip() +
  labs(title = "Comparison of the Overall Weekly Emissions for all Three Approaches and the Posmo Emissions",
       x = "User",
       y = expression("CO"["2"] ~ "Emission [g/km]")) +
  scale_x_discrete(labels=c("1af6fc8e-273d-4889-894b-5dadadd70a01" = "User 1", "6e119464-37a3-4d4e-9eaf-ef1b7e137b60" = "User 2"))
```

# Discussion

The high weekly emissions of user 1 for the distance approach mostly originate in a high use of car and train which is shown by the distance in Figure ??. Even though user 2 was more mobile than user 3 the weekly emissions were lower. This can be explained as user 2  used the car less often, which has a high emission value and user 2 often used low-emission means of transport like bus or bicycle.

The comparison of our three approaches shows that the car emissions decrease when increasing the complexity of our model. An explanation could be that we have more data points with lower velocity and only small gradients. Lower speeds were associated with lower emission values than when using a constant emission value over the entire distance. There are also many data points with only small slopes and thus the emission values are lower than for the simple approach.

*For User 2 and 3 the emissions are higher with the slope approach than for the speed approach. This may be due to the fact that they were driving on more hilly roads with probably a higher speed compared to user 1.* --> war evtl fehler, ist jetzt nicht mehr so bei mir

*Falls dir der Hintergrund nicht gefällt, hier hat es andere ;): https://leaflet-extras.github.io/leaflet-providers/preview/* Figure caption hat nicht geklappt beim rendern für leaflet

```{r plot car tracks above DHM, output = TRUE}
# convert the car data to WGS84 to use it with leaflet
posmo_car_wgs84 <- posmo_car |>
  st_transform(crs = 4326)

# create a color palette
pal <- colorFactor("Set1", posmo_car_wgs84$user_id)

# create the leaflet map with a relief as background
leaflet(posmo_car_wgs84) |>
  addTiles(urlTemplate = 'https://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}') |>
  addCircleMarkers(opacity = 0.5, 
                   color = ~pal(user_id),
                   radius = 2) |>
  addLegend(position = 'topright',
            pal = pal,
            values = ~user_id,
            title = "User ID") |> 
  addScaleBar(position = "bottomleft", scaleBarOptions(imperial = FALSE))

```

The disparities when calculating the train emissions could originate from a more restrictive approach from Posmo when calculating the emissions for trains. These differences could be explained by different assumptions in capacity utilization or the power unit of the train. Unfortunately, Posmo does not provide any information about their assumptions when calculating the emissions.

For the car emission calculations our calculations were more optimistic than the ones from Posmo. An explanation could be that our values refer to the standards of new cars registered in the last few years. These may have higher requirements to be admitted and therefore do not correspond to the average emissions. 

# Limitations

Our analysis was prone to several limitations. These ranged from inaccuracies in the data to limitations in our methods. Although only validated data from Posmo were included, there were errors resulting from the tracking where certain points were not located correctly. These lead to exceptionally high speed values or slope values. Therefore, all outliers were removed resulting in a smaller amount of data and uncertainties in the assumptions made. Another limitation was that only three different validated data sets were available which may not have well represented the general population.

Furthermore, the approaches were highly simplified and have not taken into account factors related to vehicle model or fuel type. various sources were considered for the simple approach. However, the range of the reported emission values was huge and there most likely there are other sources stating different values. The speed approach was calculated using one source (Dong et al., 2022) and extrapolating the observed values resulting in uncertainties. Similar problems occurred for the slope scenario where we only considered values from Xu et al. (2020). In addition, only positive slopes that cause an increase in emissions were considered, but no negative slopes which could have reduced potential emissions. Thus, our calculated emissions may be too high.

Furthermore, the utilization of public transport or cars was also not included. We were able to accommodate this partly because we included a range of emission values in the simple approach. Additionally, we could not take into account the effect of fuel type as the vehicle model was not known or was constantly changing.

A possible improvement would be to include changes in emissions due to acceleration and deceleration. Emissions can change significantly with changing speed (Dong et al., 2022). Figure *??* shows where there are locations with acceleration and deceleration.

```{r plot acceleration and deceleration, output = TRUE}
#| label: fig-fourteenthFigure
#| fig-cap: Posmo trajectories with acceleration and deceleration locations.
ggplot()+
  geom_sf(data = posmo, aes(color = acceleration))+
  labs(title = "Locations with acceleration",
     subtitle = "Movement trajectories in Switzerland",
     x = "E",
     y = "N")+
  guides(color = guide_legend(title = "Acceleration"))+
  coord_sf(datum = st_crs(2056))+
  theme_minimal()
```

# Conclusion 

In our analysis, we calculated emission values for different modes of transport and scenarios for emissions. We found that with increasing complexity the car emissions decrease. However, our data set consisted of only three data sets which is not representative enough to draw general conclusions. Thus, future studies should include more data sets and over longer time periods. Furthermore, additional sources should be considered to decrease uncertainties in the emission values.

# Sources

## Data

Bundesamt für Landestopographie swisstopo (no date a): DHM25. URL: <https://www.swisstopo.admin.ch/de/geodata/height/dhm25.html#download> (as at: no date) (last access: 05.06.2023).

Bundesamt für Landestopographie swisstopo (no date b): swissBOUNDARIES3D URL: <https://www.swisstopo.admin.ch/de/geodata/landscape/boundaries3d.html > (as at: no date) (last access: 05.06.2023).

Bundesamt für Statistik BFS (2023): Fahrzeuge und Transportmittelbestände des Güterverkehrs. URL: <https://www.bfs.admin.ch/bfs/de/home/statistiken/raum-umwelt/umweltindikatoren/alle-indikatoren/reaktionen-der-gesellschaft/co2-ausstoss-personenwagen.assetdetail.23908017.html> (as at: 05.04.2023) (last access: 05.06.2023).

Deutsche Physikalische Gesellschaft (2022): CO2-Emissionen beim Reisen. URL: <https://www.dpg-physik.de/vereinigungen/fachuebergreifend/ak/akjdpg/wir/arbeitsteams/nachhaltigkeit/nachgeforscht/co2-emissionen-beim-reisen> (as at: 01.01.2022) (last access: 05.06.2023).

Europäisches Parlament (2018): Senkung der Emissionen: Neue CO₂-Ziele für Pkw und leichte Nutzfahrzeug. URL: <https://www.europarl.europa.eu/news/de/headlines/society/20180920STO14027/senkung-der-emissionen-neue-co2-ziele-fur-pkw-und-leichte-nutzfahrzeuge> (as at: 15.02.2023) (last access: 05.06.2023).

Umweltbundesamt (2014): E-Rad macht mobil. Potenziale von Pedelecs und deren Umweltwirkung. URL: <https://www.umweltbundesamt.de/sites/default/files/medien/378/publikationen/hgp_e-rad_macht_mobil_-_pelelecs_4.pdf> (as at: 01.08.2014) (last access: 05.06.2023).

Umweltbundesamt (2021): Vergleich der durchschnittlichen Emissionen einzelner Verkehrsmittel Personenverkehr in Deutschland 2021. URL: <https://www.umweltbundesamt.de/sites/default/files/medien/366/bilder/dateien/uba_emissionstabelle_personenverkehr_2021_0.pdf> (as at: 01.12.2022) (last access: 05.06.2023).

Umweltbundesamt (no date): Niedrigere Geschwindigkeit spart Energie und schont die Umwelt. URL: <https://www.umweltbundesamt.at/umweltthemen/mobilitaet/mobilitaetsdaten/tempo> (as at: no date) (last access: 21.06.2023).

umweltnetz-schweiz (2019): Klimabaustein: Bus oder Bahn? URL: <https://www.umweltnetz-schweiz.ch/themen/klima/3101-klimabaustein-bus-oder-bahn.html#:~:text=Autobusse%20dagegen%20verursachen%20rund%20145,rund%2037%20g%2FPkm%20ausstossen> (as at: 18.01.2019) (last access: 05.06.2023).

World Wide Fund for Nature (2020): Autoverkehr und Elektromobilität. URL: <https://www.wwf.ch/de/unsere-ziele/autoverkehr-und-elektromobilitaet> (as at: no date) (last access: 05.06.2023).

## Literature

**Bundesamt für Statistik BFS, Bundesamt für Raumplanung ARE** (2023): Mikrozensus Mobilität und Verkehr (MZMV).

**Dong, Yaping; Xu, Jinliang; Ni, Jie** (2022): Carbon emission model of vehicles driving at fluctuating speed on highway. In: Environmental Science and Pollution Research, Vol. 30, pp. 18064-18077.

**Knupfer, G. (Handelszeitung)** (2014): Die SBB fahren im Vergleich im Schneckentempo. URL: <https://www.handelszeitung.ch/unternehmen/die-sbb-fahren-im-vergleich-im-schneckentempo-660389> (as at: 29.08.2014) (last access: 28.06.2023).

**Metcalfe, J.; Boulter & P.** (2022). Effect of speed on greenhouse gas emissions from road transport: a 
review. Report prepared by Emission Impossible Ltd, and EMM Consulting, for Waka Kotahi, 
December, 2022.

**Solaymani, Saeed** (2019): CO2 emissions patterns in 7 top carbon emitter economies: The case of transport sector. In: Energy 168, pp. 989-1001.

**Suarez, Jaime; Makridis, Michail; Anesiadou, Aikaterini; Komnos, Dimitrios; Ciuffo, Biagio; Fontaras, Georgios** (2022): Benchmarking the driver acceleration impact on vehicle energy consumption and CO2 emissions. In: Transportation Research Part D, Vol. 107.

**Wikipedia** (2022): Balmberg. URL: <https://de.wikipedia.org/wiki/Balmberg> (as at: 02.10.2022) (last access: 28.06.2023).

**Xu, Jinliang; Dong, Yaping & Yan, Menghua** (2020): A Model for Estimating Passenger-Car Carbon Emissions that Accounts for Uphill, Downhill and Flat Roads. In: Sustainability, Vol. 12, No. 2028.

